---
title: "R Notebook"
output: html_notebook
---


# Ejercicios del libro de Faraway
## 1. (Ejercicio 1 cap. 2 pág. 30)
The dataset teengamb concerns a study of teenage gambling in Britain. Fit a regression model with the expenditure on gambling as the response and the sex, status, income and verbal score as
predictors. Present the output. 
(a) What percentage of variation in the response is explained by these predictors?
(b) Which observation has the largest (positive) residual? Give the case number.
(c) Compute the mean and median of the residuals.
(d) Compute the correlation of the residuals with the fitted values.
(e) Compute the correlation of the residuals with the income.
(f) For all other predictors held constant, what would be the difference in predicted expenditure on gambling for a male compared to a female?


```{r}
pacman::p_load(faraway, broom, tidyverse)

model_1 <- 
    faraway::teengamb %>%
    lm(gamble ~ sex + status + income + verbal, data = .)


# a) El porcentaje corresponde al valor de R2    
glance(model_1)$adj.r.squared

# b) El caso cuyo residual es el mas elevado (positivo) es:
augment(model_1) %>%
    rowid_to_column() %>%
    top_n(1, .resid) %>%
    pull(rowid)

# c) la media y mediana son
augment(model_1) %>%
    summarise(
        media = mean(.resid),
        mediana = median(.resid))

# d) correlacion residuals v/s fitted.values
augment(model_1) %>%
    summarise(
        correlation = cor(.resid, .fitted)
    )

# e) correlacion residuales v/s income
augment(model_1) %>%
    summarise(
        correlation = cor(.resid, income)
    )

# f) Si todos los predictores quedan igual, se pide el poder explciativo de la variable género
faraway::teengamb %>%
    lm(gamble ~ sex, data = .) %>%
    tidy()
```


## 2. (Ejercicio 2 cap. 2 pág. 30)
The dataset uswages is drawn as a sample from the Current Population Survey in 1988. Fit a model
with weekly wages as the response and years of education and experience as predictors. Report and
give a simple interpretation to the regression coefficient for years of education. Now fit the same
model but with logged weekly wages. Give an interpretation to the regression coefficient for years
of education. Which interpretation is more natural?
```{r}
faraway::uswages %>%
    select(educ,  wage) %>%
    mutate(log_wage = log(wage)) %>%
    gather(key, value, -educ) %>%
    ggplot(aes(educ, value, color = key)) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~ key, scales = "free_y")


model_2.1 <- 
    faraway::uswages %>%
    lm(wage ~ educ + exper, data = .)

# Por cada año de educacion, aumenta el salario en 51 dolares
tidy(model_2.1)



model_2.2 <- 
    faraway::uswages %>%
    lm(log(wage) ~ educ + exper, data = .)

# Por cada año de eduacion, aumenta el salario en un 9% (beta x 100 %)
tidy(model_2.2)

```


## 3. (∗) (Ejercicio 3 cap. 2 pág. 30)
In this question, we investigate the relative merits of methods for computing the coefficients. Generate some artificial data by:
```
> x <- 1:20
> y <- x+rnorm(20)

```
Fit a polynomial in x for predicting y. Compute βˆ in two ways — by lm() and by using the direct calculation described in the chapter. At what degree of polynomial does the direct calculation method fail? (Note the need for the I() function in fitting the polynomial, that is, lm(y ~ x + I(x^2)).


## 4. (Ejercicio 4 cap. 2 pág. 30)
The dataset prostate comes from a study on 97 men with prostate cancer who were due to receive a radical prostatectomy. Fit a model with lpsa as the response and lcavol as the predictor. Record the residual standard error and the R2. Now add lweight, svi, lbph, age, lcp, pgg45 and gleason to the model one at a time. For each model record the residual standard error and the R2. Plot the trends in these two statistics.
```{r}
# esto se puede hacer como una iteracion con purrr...?
model_1 <- prostate %>%
    lm(lpsa ~ lcavol, data = .) %>% glance
model_2 <- prostate %>%
    lm(lpsa ~ lcavol + lweight, data = .) %>% glance
model_3 <- prostate %>%
    lm(lpsa ~ lcavol + lweight + svi, data = .) %>% glance
model_4 <- prostate %>%
    lm(lpsa ~ lcavol + lweight + svi + lbph, data = .) %>% glance
model_5 <- prostate %>%
    lm(lpsa ~ lcavol + lweight + svi + lbph + age, data = .) %>% glance
model_6 <- prostate %>%
    lm(lpsa ~ lcavol + lweight + svi + lbph + age + lcp, data = .) %>% glance
model_7 <- prostate %>%
    lm(lpsa ~ lcavol + lweight + svi + lbph + age + lcp + pgg45, data = .) %>% glance
model_8 <- prostate %>%
    lm(lpsa ~ lcavol + lweight + svi + lbph + age + lcp + pgg45 + gleason, data = .) %>% glance

# plot
bind_rows(model_1, model_2, model_3, model_4, model_5, model_6, model_7, model_8) %>%
    select(adj.r.squared, sigma) %>%
    rowid_to_column() %>%
    gather(key, value, -rowid) %>%
    ggplot(aes(rowid, value, color = key)) +
    geom_line()




```



## 5. (Ejercicio 5 cap. 2 pág. 30)
Using the prostate data, plot lpsa against lcavol. Fit the regressions of lpsa on lcavol and lcavol on lpsa. Display both regression lines on the plot. At what point do the two lines intersect?
```{r}

model_1 <- lm(lcavol ~ lpsa, data = prostate)
model_2 <- lm(lpsa ~ lcavol, data = prostate)

plot(prostate$lpsa, prostate$lcavol)
abline(model_1)
abline(model_2)


# calcular la intercección
# from https://stackoverflow.com/questions/7114703/finding-where-two-linear-fits-intersect-in-r

cm <- rbind(coef(model_1),coef(model_2)) # matriz de coeficientes
c(-solve(cbind(cm[,2],-1)) %*% cm[,1])


```




## 6. (Ejercicio 6 cap. 2 pág. 30)
Thirty samples of cheddar cheese were analyzed for their content of acetic acid, hydrogen sulfide and lactic acid. Each sample was tasted and scored by a panel of judges and the average taste score produced. Use the cheddar data to answer the following:
(a) Fit a regression model with taste as the response and the three chemical contents as predictors. Report the values of the regression coefficients.
(b) Compute the correlation between the fitted values and the response. Square it. Identify where this value appears in the regression output.
(c) Fit the same regression model but without an intercept term. What is the value of R2 reported in the output? Compute a more reasonable measure of the good- ness of fit for this example.
(d) Compute the regression coefficients from the original fit using the QR decomposition showing your R code.

```{r}
mod <-
   cheddar %>%
    lm(taste ~ Acetic + H2S + Lactic, data = .)

# a) 
summary(mod)



model.matrix(mod)

# b) el valor calculado corresponde a R2
augment(mod) %>%
    summarise(cor_2 = cor(.fitted, taste)^2) %>%
    pull(cor_2)


# c) 
cheddar %>%
    lm(taste ~ Acetic + H2S + Lactic -1, data =.) %>%
    glance %>%
    pull(r.squared)


```


## 7. (Ejercicio 7 cap. 2 pág. 31)
An experiment was conducted to determine the effect of four factors on the resistivity of a semiconductor wafer. The data is found in wafer where each of the four factors is coded as − or + depending on whether the low or the high setting for that factor was used. Fit the linear model resist ~ x1 + x2 + x3 + x4.
(a) Extract the X matrix using the model.matrix function. Examine this to determine how the low and high levels have been coded in the model.
(b) Compute the correlation in the X matrix. Why are there some missing values in the matrix?
(c) What difference in resistance is expected when moving from the low to the high level of x1?
(d) Refit the model without x4 and examine the regression coefficients and standard errors? What stayed the the same as the original fit and what changed?
(e) Explain how the change in the regression coefficients is related to the correlation matrix of X.

```{r}
mod <- 
wafer %>%
    lm(resist ~ x1 + x2 + x3 + x4, data =.)


model.matrix(mod)

wafer
```



## 8. (∗) (Ejercicio 8 cap. 2 pág. 31)
An experiment was conducted to examine factors that might affect the height of leaf springs in the suspension of trucks. The data may be found in truck. The five factors in the experiment are set
to − and + but it will be more convenient for us to use −1 and +1. This can be achieved for the first factor by:
```
truck$B <- sapply(truck$B, function(x) ifelse(x == "-",-1,1))
```
Repeat for the other four factors.
(a) Fit a linear model for the height in terms of the five factors. Report on the value of the regression coefficients.
(b) Fit a linear model using just factors B, C, D and E and report the coefficients. How do these compare to the previous question? Show how we could have anticipated this result by examining
the X matrix.
(c) Construct a new predictor called A which is set to B+C+D+E. Fit a linear model with the predictors A, B, C, D, E and O. Do coefficients for all six predictors appear in the regression
summary? Explain.
(d) Extract the model matrix X from the previous model. Attempt to compute βˆ from (XT X)
−1XT y. What went wrong and why?
(e) Use the QR decomposition method as seen in Section 2.7 to compute βˆ. Are the results
satisfactory?
(f) Use the function qr.coef to correctly compute βˆ.



# Ejercicios del libro de Carmona
## 1. (Ejercicio 2.1 del Capítulo 2 página 41)
Una variable Y toma los valores y1, y2 y y3 en función de otra variable X con los valores x1, x2 y x3. Determinar cuales de los siguientes modelos son lineales y encontrar, en su caso, la matriz de diseño para x1 = 1, x2 = 2 y x3 = 3.
a) yi = β0 + β1xi + β2(x2i − 1) + i
b) yi = β0 + β1xi + β2e xi + i
c) yi = β1xi(β2tang(xi)) + i


## 2. (Ejercicio 2.4 del Capítulo 2 página 42)
Cuatro objetos cuyos pesos exactos son β1, β2, β3 y β4 han sido pesados en una balanza de platillos de acuerdo con el siguiente esquema:
β1 β2 β3 β4 peso
1 1 1 1 9.2
1 −1 1 1 8.3
1 0 0 1 5.4
1 0 0 −1 −1.6
1 0 1 1 8.7
1 1 −1 1 3.5
Hallar las estimaciones de cada βi y de la varianza del error.

## 3. (∗) (Ejercicio 3.2 del Capítulo 3 página 54)
En un modelo lineal, la matriz de diseño es


1 1 1 1 1
1 0 1 0 0
1 1 1 0 0
1 0 1 1 1


Hallar la expresión general de las funciones paramétricas estimables.

# 4. (∗) (Ejercicio 3.7 del Capítulo 3 página 55)
Consideremos el modelo lineal
y1 = β1 + β2 + 1
y2 = β1 + β3 + 2
y3 = β1 + β2 + 3
Se pide:
1) ¿Es la función paramétrica
ψ = β1 + β2 + β3
estimable?
2) Probar que toda función paramétrica
ψ = a1β1 + a2β2 + a3β3
es estimable si y sólo si a1 = a2 + a3.

# 5. (∗) (Ejercicio 3.8 del Capítulo 3 página 56)
Consideremos el modelo lineal
y1 = µ + α1 + β1 + 1
y2 = µ + α1 + β2 + 2
y3 = µ + α2 + β1 + 3
y4 = µ + α2 + β2 + 4
y5 = µ + α3 + β1 + 5
y6 = µ + α3 + β2 + 6
(a) ¿Cuando es λ0µ + λ1α1 + λ2α2 + λ3α3 + λ4β1 + λ5β2 estimable?
(b) ¿Es α1 + α2 estimable?
(c) ¿Es β1 − β2 estimable?
(d) ¿Es µ + α1 estimable?
(e) ¿Es 6µ + 2α1 + 2α2 + 2α3 + 3β1 + 3β2 estimable?
(f) ¿Es α1 − 2α2 + α3 estimable?
(g) Hallar la covarianza entre los estimadores lineales MC de las funciones paramétricas β1 − β2 y
α1 − α2, si éstas son estimables.
(h) Hallar la dimensión del espacio paramétrico.
(i) Obtener una expresión del espacio de los errores.

# 6. (∗) (Ejercicio 3.10 del Capítulo 3 página 56)
Un transportista realiza diversos trayectos entre tres poblaciones A, B y C. En cuatro dias consecutivos ha hecho los recorridos que muestra la siguiente tabla:
trayecto km
A → B → A → C 533
C → A → C → B 583
B → C → A → C → A → B → A 1111
A → B → A → C → A → B → A 1069
donde el kilometraje es, por diversas causas, aproximado.
(a) Proponer un modelo lineal, con la matriz de diseño y las hipótesis necesarias, para estimar las
distancias kilométricas entre las tres poblaciones.
Con los datos proporcionados, ¿es posible estimar las distancias entre las tres poblaciones?
¿Cuales son las distancias o funciones paramétricas estimables (fpe) en este modelo?
(b) ¿Se puede estimar el kilometraje del trayecto MBC → B → A → C → MAC , donde MIJ es
el punto medio entre dos poblaciones? ¿Es una buena estimación? ¿Cual es el error de esta
estimación?
