---
title: "Untitled"
author: "Alejandro Keymer"
date: "3/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(faraway)
```


# Ejercicios del libro de Faraway
## 1. (Ejercicio 1 cap. 14 pág. 220)
Using the `teengamb` data, model `gamble` as the response and the other variables as predictors, investigate the possibility of interactions between sex and the other predictors. Interpret your final model.

```{r}
df <- 
    teengamb  %>%
    mutate(sex = factor(sex, labels = c('M', 'F')))
    

mod <- lm(gamble ~ (status + income + verbal) * sex, data = teengamb)

summary(mod)


broom::augment(mod) %>%
    ggplot(aes(gamble, income, color = factor(sex))) +
    geom_point()

broom::augment(mod) %>%
    ggplot(aes(.fitted, .resid, color = factor(sex))) +
    geom_point()
```





## 2. (Ejercicio 2 cap. 14 pág. 221)
Using the `infmort` data, find a simple model for the infant mortality in terms of the other variables. Be alert for transformations and unusual points. Interpret your model by explaining what the regression parameter estimates mean.
```{r}
mod <- step(lm(mortality ~ ., infmort))
summary(mod)
```



## 3. (Ejercicio 3 cap. 14 pág. 221)
Plot the ToothGrowth data with len as the response. Fit a linear model to the data and check for possible simplification. Display the fitted regression lines on top of the data.

## 4. (*) (Ejercicio 4 cap. 14 pág. 221)
Investigate whether the side variable can be used in a model for all the chredlin data rather than as a subsetting variable as in the analysis in the text.

## 5. (*) (Ejercicio 5 cap. 14 pág. 221)
Find a good model for the uswages data with wages as the response and all the other variables as predictors.

6. (Ejercicio 6 cap. 14 pág. 221)
The dataset clot contains the clotting times of blood varying as percentage concentration of
prothrombin-free plasma. There are two lots of thromboplastin.
(a) Plot the data using a different plotting symbol according to the lot.
(b) Find the transformation of the two continuous variables to form a linear relationship.
(c) Does the time to clot vary according to concentration differently in the two lots?
(d) Check the assumptions of your model using regression diagnostics.
(e) At what percentage concentration is the predicted time the same for the two lots?
7. () (Ejercicio 7 cap. 14 pág. 221)
The wealth in billions of dollars for 232 billionaires is given in fortune.
(a) Plot the wealth as a function of age using a different plotting symbol for the different regions
of the world.
(b) Plot the wealth as a function of age with a seperate panel for each region.
(c) Determine a transformation on the response to facilitate linear modeling.
(d) What is the relationship of age and region to wealth?
(e) Check the assumptions of your model using appropriate diagnostics.
8. (Ejercicio 1 cap. 15 pág. 233)
Using the pulp data, determine whether there are any differences between the operators. What is
the nature of these differences?
9. (Ejercicio 2 cap. 15 pág. 233)
Determine whether there are differences in the weights of chickens according to their feed in the
chickwts data. Perform all necessary model diagnostics.
10. (Ejercicio 3 cap. 15 pág. 233)
Using the PlantGrowth data, determine whether there are any differences between the groups.
What is the nature of these differences? Test for a difference between the average of the two
treatments and the control.
11. (Ejercicio 4 cap. 15 pág. 233)
Using the infmort data, perform a one-way ANOVA with income as the response and region as
the predictor. Which pairs of regions are different? Now check for a good transformation on the
response and repeat the comparison.
12. () (Ejercicio 5 cap. 15 pág. 233)
The anaesthetic data provides the time to restart breathing unassisted in recovering from general
anaesthetic for four treatment groups.
(a) Produce a boxplot depicting the data. Comment on any features of interest.
(b) Make an appropriate stripchart of the data.
(c) Produce versions of the previous two plots using the ggplot2 package. Show how you can
overlay these two plots.
(d) Fit a one-factor model for the recovery times and test for a difference between the two groups.
(e) Try the Box-Cox transformation method. Explain what went wrong.
(f) Try a square root transformation on the response. Are the diagnostics satisfactory? Is there
a significant difference among the treatment groups?
13. () (Ejercicio 6 cap. 15 pág. 233)
Data on the butterfat content of milk from Canadian cows of five different breeds can be found in
the butterfat dataset. Consider only mature cows.
(a) Plot the data and interpret what you see.
(b) Test for a difference between the breeds.
(c) Check the diagnostics for your model.
(d) Produce a plot to check for differences in butterfat between breeds. Which pairwise differences
are not statistically signifcant?
Página 2 de 2